Sophize Mathematics Library

### 2022-05-27 math_library/sophize_math.tex
CICM '22 1st submitted version

\subsection*{The Sophize Platform}

Sophize is a novel mathematics library and discussion platform. The author has developed the platform over the course of the last two years at \url{https://sophize.org}. Sophize's primary mission is to help users find existing proofs of mathematical statements, to discover new proofs, and to utilize this knowledge in their work. We combine knowledge from multiple resources and have accumulated thousands of definitions, theorems, and proofs from a wide variety of sources, including some published research, the PlanetMath encyclopedia, Wikipedia, and the Metamath formal system \cite{metamath}.

\subsection*{Motivation}

In the past couple of decades, the Internet has revolutionized our life by giving us access to seemingly unlimited information. Information from a wide variety of topics ranging from human medicine to rocket science is now just a few keystrokes away. However, even with this information we seem to having trouble understanding the complex world around us. In fact, dealing with the vast amounts of information thrown at us seems to be one of the primary problems of our generation. While we have access to a wide range of opinions on almost every topic imaginable, it has becoming increasingly difficult to decide which one to trust. It seems that you can find "proofs" for anything and everything we would like to believe.

Thus the author's primary motivation for this project is to create a library that not just lays out all the information of the World but organizes in a way that help us make better sense of it. A library that helps us access justifications from a variety of viewpoints and help us collaboratively and objectively evaluate issues in the justifications or the viewpoints they arise from. Clearly, this is challenging goal.

We begin our journey with a focus on pure Mathematics, where knowledge is objectively less complex as compared to other fields. In mathematics, only logically sound arguments are acceptable as justifications. These arguments develop from different foundational theories and logical consistency is the primary criteria of correctness for them. Thus, as a Mathematics library, Sophize's goal is to help our users access to mathematical proofs (chains of arguments) from a variety of foundational theories (extended to `belief sets' as described in Section \ref{sec:bset}) and help us collaboratively and objectively evaluate issues in the argument or the belief sets they arise from.

With our current scheme we can incorporate almost all mathematics. A natural extension would be model observational data and develop criteria to quantify empirical correspondence in order to have the ability to incorporate scientific knowledge. But, for now we are focused on building a state-of-the-art Mathematics library that is valuable to our community.

\subsection*{Proofs in Sophize}

\begin{figure}[!ht]
\begin{center}
\fbox{\includegraphics[width=\textwidth]{proof_tree}}
\caption{A proof displayed as a graph (tree) of arguments and propositions on the Sophize platform.}
\label{proof_tree}
\end{center}
\end{figure}

@@

This work can also be seen as a step towards formalizing the network of information that exists in the connections of mathematical objects. The committee on planning a global library of the mathematical sciences recognized that this network is largely unexplored, and formalizing it has tremendous potential to accelerate math research \cite{sciences2014developing}.

Hence, we required novel knowledge organization techniques to connect mathematical entities from such a wide range of formal and informal knowledge sources. And thus, one of the main challenges towards building the Sophize platform was to create a simple way to embed mathematical entities on the web.

\subsection*{Contribution}
In this paper, we present the novel concepts and features of Sophize Mathematics Library. First, we expound Sophize's scheme to build a trust-focused logical network of mathematical knowledge. In section \ref{sec:pgm}, we describe Sophize's proof-generating machines which are computer programs that generate proofs that integrate with Sophize's knowledge graphs.

@@

Section \ref{sec:lib} discusses some other challenges and contributions made towards making Sophize a modern mathematics library. Finally, Section \ref{sec:summary} concludes the paper.

\section{Sophize's Knowledge Organization Scheme}
At its core, Sophize's approach is to simply keep track of all arguments in favour or against any proposition. Use of some form of arguments is the common factor amongst all mathematical (and perhaps all scientific) knowledge. We use this fact to unify knowledge from multiple sources in a meaningful way as described below.

We use the set of all known arguments to track knowledge from a variety of foundational theories. For theoretical knowledge, i.e., knowledge that doesn't involve empirical data, the primary criteria of validity is internal (logical) consistency. Thus, for each foundational theory (`belief set' described below), we also keep track of any contradictions that arise from those foundations.

\subsection{Core Concepts}
Sophize uses the following concepts to logically organize theoretical knowledge.

\paragraph{A \textbf{resource} is an abstract concept inherited by all other top-level concepts like terms, propositions, arguments, and belief sets. Each resource has a URI and contains fields such as search tags and citations.}

A \textbf{URI} consists of two parts: a namespace-like identifier called its dataset-id, which indicates the data source, and a resource-id that specifies the resource type and its unique name in the data source. The dataset-id may be omitted if it can be inferred from the surrounding context.

@@

\paragraph{An \textbf{argument} is a set of propositions called premises along with a concluding proposition that is claimed to follow from the premises. In addition, most arguments include supporting text that explains how the conclusion follows from the premises. A proof is seen here as a directed graph of arguments and propositions.}

\paragraph{A \textbf{belief set} is roughly the set of axiomatic propositions that make up the foundations of a theory. Belief sets will be described in more detail in subsection \ref{sec:bset}.}

\paragraph{Loosely defined, a \textbf{proof-generating machine} (PGM) is a computer program that generates the proof for an input proposition. PGMs are described in more detail in section \ref{sec:pgm}.}

\subsection{Argument Graph}
An argument can be seen as a simple graph (see Fig.~\ref{argument}) with two types of nodes - (a) a single node representing the argument itself and (b) a set of proposition nodes representing the argument's conclusion and premises. The edges of this graph are directed and go from (a) premise nodes to argument node and (b) argument node to conclusion node.

\begin{figure}[!ht]
\begin{center}
\fbox{\includegraphics[height=3cm]{argument}}
\caption{An argument represented as a graph.}
\label{argument}
\end{center}
\end{figure}

A set of arguments can thus be represented as a single graph that we call an \textbf{argument graph} (Fig.~\ref{argument_graph}).

\begin{figure}[!ht]
\begin{center}
\fbox{\includegraphics[height=5.5cm]{argument_graph}}
\caption{A set of arguments forming an argument graph.}
\label{argument_graph}
\end{center}
\end{figure}

\subsection{Generating Proof Graphs}
An argument graph along with a set of set of axioms results in a \textbf{proof graph}. Any reachability algorithm starting with the axiom nodes is sufficient to create the proof graph (See Fig.~\ref{proof_graph1}).

\begin{figure}[!ht] 
\begin{center}
\fbox{\includegraphics[height=5.5cm]{proof_graph1}}
\caption{Proof graph from argument graph (Fig.~\ref{argument_graph}) using `Prop 2', `Prop 3', Prop 4' are axioms.}
\label{proof_graph1}
\end{center}
\end{figure}

A proof graph is a directed acyclic graph and contains complete proofs (including all alternate proofs) for all its propositions.

\begin{figure}[!ht]
\begin{center}
\fbox{\includegraphics[height=4cm]{proof_prop7}}
\caption{Proof of `Prop 7' from proof graph in Fig.~\ref{proof_graph1} uses arguments `Arg 4' and `Arg 5'.}
\label{proof_prop7}
\end{center}
\end{figure}

Note that an argument graph combines with different sets of axioms to generate different proof graphs. The same argument may be utilized in multiple belief sets (Fig.~\ref{proof_graph2}).

\begin{figure}[!ht]
\begin{center}
\fbox{\includegraphics[height=5.5cm]{proof_graph2}}
\caption{Proof graph from the same argument graph (Fig.~\ref{argument_graph}) using `Prop 1', `Prop 2', and Prop 7' as axioms.}
\label{proof_graph2}
\end{center}
\end{figure}

@@

The correctness of proofs obtained by the above process is contingent on the validity of the arguments in the argument graph. The validity of an argument is, in principle, independent of other arguments and objectively verifiable. Formally defined logistic systems indeed have a mechanical procedure to independently verify validity of arguments.

However, only a small section of Mathematics is developed using formally verified arguments. Vast majority of mathematics is verified via peer review, some of this knowledge is curated by experts in books, encyclopaedias. The claim of validity of proofs (arguments) is judged based on the various parameters such as academic standing of the claimant, reputation of the peer reviewing journal, the history of errors of the book or encyclopaedia containing them, etc. Clearly, these criteria are subjective and each individual's criteria of accepting validity of an argument can vary to some extent.

@@

\begin{figure}[!ht]
\begin{center}
\fbox{\includegraphics[height=4cm]{validity_criterion}}
\caption{Validity criteria as an argument premise.}
\label{validity_criterion}
\end{center}
\end{figure}

To easily manage validity criteria, the relationships between validity criteria are also maintained in the argument graph using argument-like dependency graphs. Eg. "Published in journal with SJR score $>$ 0.67" $\rightarrow$ "Published in Annals of Mathematics". By maintaining relationships like above, Sophize allows its users to conveniently choose multiple sources they consider to be reliable with a single manually chosen criterion.

@@

A proposition is true if it is the conclusion of a valid argument and the premises of the argument are also true. Of course, this means that we need to start with some propositions that are assumed to be true without further justification. Traditionally, these are mathematical axioms and are considered to self-evidently true. In mathematical inquiry, theorems are proven/disproven in one or more multiple theories, each with its own set of axioms. Each theory may have its theoretical advantages and practical uses and there are no universal basis to choose one theory over another when trying to assert a proposition's truth/falsity.

Thus, Sophize allows any set of propositions to be considered true without evidence. Such a set of propositions is called a \textbf{belief set}. Propositions representing argument validity criteria are also part of belief sets. All propositions that are considered to be true in a belief set without proof are called its `unsupported propositions'. In addition, a belief set can also contain a number of `proof-generating machines' (e.g. for managing axiom-schemas) that will be discussed later. For ease of use, belief sets can also contain other belief sets.

The truth-value of a proposition is only defined within a belief set - there is no notion of of absolute or universal truth. Sophize creates proof graphs for all its belief sets.

@@

\subsection{Advantages and Limitations}

Efficiently modelling different axiomatic systems that have a large overlap of arguments(eg. euclidean/parabolic/hyperbolic geometry, ZF/ZFC etc)

Conjectural knowledge (Riemann hypothesis). 
Practical utility of conjectural knowledge - eg. a lot of information security systems assume "P is not NP"
Proof tailored to needs. Proof graph can be easily converted to prose too.

\section{Proof-Generating Machines (PGM)}
\label{sec:pgm}
A lot of mathematical knowledge is computed. Computation machines (e.g. calculators, algebras) can be seen as performing the following tasks
\begin{itemize}
\item Parsing the input into a language that can represent math formulae. 
\item Finding and outputting an equivalent form of the formula.
\end{itemize}
Thus, their focus is not on finding the proof but only finding the equivalent (simplified) representation of the input formula (Although many computer algebras do show the steps of the computation.). With Sophize, the focus instead is to generate proofs of the input propositions. A \textbf{Proof-generating machine} (PGM for short) is a computer program that generates a partial proof graph against or in favour of an argument. PGMs can also have the capability to perform tasks 1 and 2, mentioned above.

@@

The Sophize platform itself merely forwards the request to the associated PGM and performs some checks on the output returned. Any http server that can process a proof request can be easily plugged in to Sophize platform using its HTTP address. One server can implement multiple PGMs.

\subsection{`Active' PGMs}
As with other parts of Sophize, the validity of the output of a PGM is not taken for granted. Each PGM must specify all the propositions (premises) that it will use in the root nodes of the proof graph it returns. In addition, a PGM may also delegate parts of proofs to another PGM. These PGMs must also be specified.

A PGM is considered `active' in a belief set iff:
\begin{enumerate}
\item all of its premises are true in the belief set.
\item all the PGMs it delegates-to are active.
\end{enumerate}

If the PGM returns a proof graph that utilizes As with an argument, one or more premises is used to indicate the validity of the PGM (Eg. tested by X, verified by some committee, another program etc). Thus, to be considered active in a belief set, the PGM must have a verification criteria that matches the belief set's requirements. The output of a PGM is not used in belief sets in which it is not active.

@@

\item Truth value, i.e., whether the PGM considers the input proposition (provided or computed) true or false. PGM can also indicate that it doesn't understand the input or prove/disprove the proposition.

@@

\begin{enumerate}
\item Arguments whose the premises are limited to the premises of the PGM.
\item Special `Arguments' that indicate that the generation of the remaining proof must be delegated to another PGM. In case of large proofs, a PGM may also delegate to itself after generating some (say 100) arguments.
\end{enumerate}
\subsection{Materializing results}

@@

In the author's view, it is crucial to maintain the proofs graphs for building trustable knowledge. However, the presentation of proofs as graphs to users is perhaps not very convenient. Traditional text based narration remains indispensable when presenting knowledge to the reader.

@@

With Sophize Markdown, we can easily add \LaTeX\space by enclosing it between two \$ signs (or \$\$ for display math). The detailed specification of Sophize Markdown is available here. Many of its features are demonstrated at: \url{https://youtu.be/5UYOpQwcjCk}. We highlight the prominent features here.

\subsection{Embedding semantic data}

With Sophize Markdown we can easily link resources such as terms, propositions, and arguments. Links can be added using the pound sign (\#) and the resource's URI. There are multiple resource link options that can be used for changing the behavior and text of the resource link. The following examples demonstrate these options.

\subsubsection*{Example 1}
\begin{verbatim}
#P_cosine_law|UC generalizes the #P_pythagorean for all kinds of #(T_triangle, 'triangles').
\end{verbatim}

\paragraph{The above code is rendered as:}
\begin{mdframed}
\dashuline{Law of cosines} generalizes the \dashuline{Pythagorean theorem} for all kinds of \dashuline{triangles}.
\end{mdframed}

The link options 'UC/LC' makes the link text upper-cased or lower-cased.

\subsubsection*{Example 2}
\begin{verbatim}
#T_matrix_multiplication|NO_LINK is a #(T_commutative_property, 'non-commutative'|NAV_LINK)
#T_binary_operation.
\end{verbatim}

\paragraph{The above code is rendered as:}
\begin{mdframed}
Matrix multiplication is a \underline{non-commutative} \dashuline{binary operation.}
\end{mdframed}
This example shows the various link behaviours. The first element shows `\#wiki/T\_matrix\_multiplication' on mouse-hover. The second link navigates the page to a different URL. Clicking on the third link pops up a modal dialog box like so:

\begin{figure}
\begin{center}
\fbox{\includegraphics[width=\textwidth]{modal}}
\caption{Modal dialog box for the term binary operation. These boxes can also have content that uses all three types of links. }
\label{modal}
\end{center}
\end{figure}

\subsubsection*{Example 3}

@@

A conic section (or simply conic) is a curve obtained as the intersection of the
surface of a #wiki/T_cone with a #wiki/T_plane. There are three types of conic sections.

@@

\paragraph{The above code is rendered as:}
\begin{mdframed}
\section*{Conic Section}
A conic section (or simply conic) is a curve obtained as the intersection of the surface of a \dashuline{cone} with a \dashuline{plane}. There are three types of conic sections.

@@

In the above example, the `EXPAND' option expanded the definition of the term, and statement of the proposition in place. Thus we were able to easily re-use concepts and theorems already extracted from other sources (say, Wikipedia and the Oxford dictionary in this case).

@@

Sophize markdown adds an icon next to each proposition link. This icon indicates whether the proposition has been proven or dis-proven (or both) in the currently selected belief set using a check-mark or a cross. Clicking on the icon brings up the proof graph as shown in Fig.~\ref{proof_tree}. If the users switches from one belief set to another, the icon is updated based on the truth values in the new belief set but other parts of the article are unchanged.

Similarly, an icon is added next to a proof generating machine which indicates whether the PGM is active or not.

\subsection{Formal Language Support}

In formal languages, definitions of terms, statements of propositions, and supporting argument text may be parse-able by an external parser. We can convert the externally parsed output into Markdown in such a case, where each term is automatically linked to the appropriate resource. This provides a convenient interface, where the user types in the native language, and the final output automatically allows users to explore all concepts that make up the input statement in depth. Currently, this is demonstrated in the use of Sophize Markdown with the Metamath language.

\begin{figure}
\begin{center}
\fbox{\includegraphics[height=2.25cm]{formal}}
\caption{The prime number theorem written in Metamath language and rendered as Markdown in real time.}
\label{formal}
\end{center}
\end{figure}

\section{Building a Modern Mathematics Library}
\label{sec:lib}
The argument and proof graphs in help us build a network of information that spans across a wide variety of sources (including live results from PGMs). Using different belief sets, users can view mathematical knowledge arising from different theoretical foundations and choose what they consider to be reliable. With Sophize Markdown, users can easily embed semantic information, thus making it convenient to explore the network of knowledge that is being built. However, a few more practical issues need to be overcome to build a modern scalable library with reliable information.

\subsection{Semantic Data Extraction}
Utilizing the full range of features offered by the Sophize platform requires the availability of semantic information. However, extracting semantic information from unstructured or semi-structured literature is a difficult problem. This problem can be seen as composed of several sub-problems:
\begin{itemize}
\item Finding definitions and propositions in prose.
\item Associating the appropriate entity with a term used in prose for quick lookup.
\item Identifying arguments and along with premises and conclusion.
\item \label{prob:dup} Identifying duplicate definitions and propositions used in different sources.
\end{itemize}

Solving these problems would need manual effort (e.g. crowd-sourcing) combined with automated extraction tools. Recent progress in machine learning and natural language processing techniques does make this problem seem tractable.

\subsubsection{Connecting Across Datasets}
The last sub-problem mentioned above is crucial to help make connections across datasets. For example, once we have identified that the `Jensen's Inequality' in a published paper is same as the Planetmath proposition with the same name, Planetmath's proof is automatically made available to the reader of the published paper.

However, figuring out if two sources are referring to the same proposition or a different version of the same proposition is often unclear. When working with formal systems using different languages, another related problem arises. There is no general way to claim equivalence of propositions across different formal systems. There are some projects, like Formal Harmony and theory morphisms, that are trying to understand in what sense these propositions are same and how we can translate one to another. But these projects are still in early stages and do not provide a general solution.

However, we can utilize curated translations, such as those available from Logipedia, by representing them using arguments. The conclusion of such an argument is the proposition in target language and the premise would be the proposition in source language. For bi-directional translations two such arguments will be required. Such translations do not stand on a very firm foundation and that should be indicated in the validation criteria of the argument. A user can choose not to accept such criteria. In that case, these translations will not be used in the generating proof graphs. Thus we can create proof graphs that have arguments from different languages.

There are some cases where translations occur on firmer grounds, e.g., use of theory morphisms or translations of HOL-light proofs to Metamath. The author has not studied these cases in detail but expects that it to be possible to formulate such cross-language translations using the argument graphs.

@@

Sophize divides its data into groups called datasets for copyright management and data access rules. For example, the 'wiki' dataset that extracts knowledge from Wikipedia mathematics glossaries releases its data under `CC-BY-SA' license, whereas the `metamath' dataset is public (`CC0'). Our signed-in users are free to create their own datasets and we allow them to choose who can read, write or comment in these datasets. Another interesting access control is the restriction of use of resources (especially argument-validity criteria) outside an otherwise open-to-read dataset.

Information can be added to Sophize platform using intuitive web forms. However, it is infeasible to extract information from large repositories in this manner. Sophize publishes its data schema and helper libraries (explained in Section \ref{sec:core}) in multiple languages. The extracted information is can thus be stored as text (JSON) files in a repository. Sophize provides dataset owners the ability to sync knowledge such repositories to the central Sophize database.

>>>>

\subsection{The Sophize Platform}

Sophize is a novel mathematics library and discussion platform. The author has developed the
platform over the course of the last few years at \url{https://sophize.org}. Sophize's
primary mission is to help users find existing proofs of mathematical statements, to
discover new proofs, and to utilize this knowledge in their work. We combine knowledge from
multiple resources and have accumulated thousands of definitions, theorems, and proofs from
a wide variety of sources, including some published research, the PlanetMath encyclopedia,
Wikipedia, and the Metamath formal system \cite{metamath}.

\subsection{Motivation}

In the past couple of decades, the Internet has revolutionized our life by giving us access
to seemingly unlimited information. Information from a wide variety of topics ranging from
human medicine to rocket science is now just a few keystrokes away. However, it seems
that we are having even more trouble making sense of the World around us. In fact, dealing with the vast amounts of information thrown at us seems to be one of the most important problems of our generation. While we have access to a wide range of opinions on almost every topic imaginable, it has become increasingly difficult to decide which one to trust. It
seems that we can find "proofs" for anything and everything we would like to believe.

Thus the author's primary motivation for this project is to create a library that not just
lays out all the information about the World but organizes it in a way that helps us make
better sense of it. A library that helps us access justifications from a variety of
viewpoints and allows us to collaboratively and objectively evaluate issues in the
justifications or the viewpoints they arise from. Clearly, this is a challenging goal.

We begin our journey with a focus on pure Mathematics, where knowledge is objectively less
complex as compared to other fields. In mathematics, only logically sound arguments are
acceptable as justifications. These arguments develop in different foundational theories;
and logical consistency is the primary criterion of correctness. Thus, as a
Mathematics library, Sophize's goal is to help our users access mathematical proofs (arguments) from a variety of belief sets (roughly - foundational theories along with indicators of trust in knowledge sources) and allow our users to collaboratively and objectively evaluate issues in the arguments or the belief sets they developed in.

With our current scheme, we can incorporate nearly all mathematics. A natural extension
would be to model observational data and develop criteria to quantify empirical
correspondence in order to have the ability to incorporate scientific knowledge. But, for
now, we are focused on developing a state-of-the-art library for the Mathematics community.

\subsection{Proofs in Sophize}
\label{sec:sophize_proofs}

\begin{figure}[!ht]
\begin{center}
\fbox{\includegraphics[height=8.25cm]{proof_tree}}
\caption{A proof displayed as a graph (tree) of arguments and propositions on the Sophize platform.}
\label{proof_tree}
\end{center}
\end{figure}

@@

This work can also be seen as a step towards formalizing the network of information that exists in the connections of mathematical objects. The committee on planning a global library of the mathematical sciences recognized that this network is largely unexplored, and formalizing it has tremendous potential to accelerate math research \cite{sciences2014developing}. Hence, we required novel techniques to connect mathematical entities from such a wide range of formal and informal knowledge sources.

\subsection{Contribution}
In this paper, we present the novel concepts and features of Sophize Mathematics Library. First, we expound Sophize's scheme to build a trust-focused logical network of mathematical knowledge. In section \ref{sec:pgm}, we describe Sophize's proof-generating machines which are computer programs that generate proofs that integrate with Sophize's knowledge graphs.

@@

Section \ref{sec:lib} discusses some other challenges and contributions made towards making Sophize a modern mathematics library. Finally, Section \ref{sec:conclusion} concludes the paper.

\section{Sophize's Knowledge Organization Scheme}
At its core, Sophize's approach is to simply keep track of all arguments in favour or against any proposition. The use of some form of arguments is the common factor amongst all mathematical (and perhaps most scientific) knowledge. We use this fact to unify knowledge from multiple sources in a meaningful way as described below.

We use the set of all known arguments to track knowledge from a variety of foundational theories. For theoretical knowledge, i.e., the knowledge that doesn't involve empirical data, the primary criterion of validity is internal (logical) consistency. Thus, for each foundational theory (`belief set' described below), we also keep track of any contradictions that arise from those foundations.

\subsection{Core Concepts}
\label{sec:core}
Sophize uses the following concepts to logically organize theoretical knowledge. The datamodel for these concepts is published in JSON schema \cite{sophize_datamodel}.

\paragraph{A \textbf{resource} is an abstract concept inherited by all other top-level concepts like terms, propositions, arguments, and belief sets. Each resource has a URI and contains fields such as search tags and citations.}
\paragraph{}
A \textbf{URI} consists of two parts: a namespace-like identifier called its dataset-id, which indicates the data source, and a resource-id that specifies the resource type and its unique name in the data source. The dataset-id may be omitted if it can be inferred from the surrounding context.

@@

\paragraph{An \textbf{argument} is a set of propositions called premises along with a concluding proposition that is claimed to follow from the premises. In addition, most arguments include supporting text that explains how the conclusion follows from the premises.}

\paragraph{A \textbf{belief set} is roughly the set of axiomatic propositions that make up the foundations of a theory. Belief sets will be described in subsection \ref{sec:bset}.}

\paragraph{Loosely defined, a \textbf{proof-generating machine} (PGM) is a computer program that generates proofs for an input proposition on-the-fly. PGMs are described in detail in section \ref{sec:pgm}.}

\subsection{Argument Graph}
An argument can be seen as a simple graph (see Fig.~\ref{argument_graph}) with two types of nodes - (a) a single node representing the argument itself and (b) a set of proposition nodes representing the argument's conclusion and premises. The edges of this graph are directed and go from (a) premise nodes to argument node and (b) argument node to conclusion node. A set of arguments can thus be represented as a single graph that we call an \textbf{argument graph}.

\begin{figure}[!ht]
\begin{center}
\fbox{\includegraphics[width=\textwidth]{argument_graph}}
\caption{An argument represented as a graph (left). A set of arguments forming an argument graph (right).}
\label{argument_graph}
\end{center}
\end{figure}

\subsection{Generating Proof Graphs}
A \textbf{proof graph} is a directed acyclic graph and contains complete proofs (including all alternate proofs) for all its propositions. A proof graph can be generated from an argument graph by starting with the axiom nodes and finding all nodes that can be reached from them (See Fig.~\ref{proof_graph1}).

\begin{figure}[!ht] 
\begin{center}
\fbox{\includegraphics[height=5cm]{proof_graph1}}
\caption{Proof graph from argument graph (Fig.~\ref{argument_graph}) using `Prop 2', `Prop 3', Prop 4' as axioms.}
\label{proof_graph1}
\end{center}
\end{figure}

\begin{figure}[!ht]
\begin{center}
\fbox{\includegraphics[height=4cm]{proof_prop7}}
\caption{Proof of `Prop 7' from proof graph in Fig.~\ref{proof_graph1} uses `Arg 4' and `Arg 5'.}
\label{proof_prop7}
\end{center}
\end{figure}

Note that an argument graph combines with different sets of axioms to generate different proof graphs. The same argument may be utilized in multiple proof graphs (Fig.~\ref{proof_graph2}).

\begin{figure}[!ht]
\begin{center}
\fbox{\includegraphics[height=5cm]{proof_graph2}}
\caption{Proof graph from the same argument graph (Fig.~\ref{argument_graph}) using `Prop 1', `Prop 2', and Prop 7' as axioms.}
\label{proof_graph2}
\end{center}
\end{figure}

@@

The correctness of proofs obtained by the above process is contingent on the validity of the arguments in the argument graph. The validity of an argument is, in principle, independent of other arguments and objectively verifiable. Formally defined logistic systems indeed have a mechanical procedure to independently verify the validity of arguments.

However, only a small section of Mathematics is developed using formally verified arguments. The vast majority of mathematics is verified via peer review, and some of this knowledge is curated by experts in books, encyclopaedias. The claim of validity of proofs (arguments) is judged based on various parameters such as the academic standing of the claimant, the reputation of the peer reviewing journal, the history of errors in the book or encyclopaedia containing them, etc. Clearly, these criteria are subjective and each individual's criteria for accepting the validity of an argument can vary to some extent.

@@

\begin{figure}[!ht]
\begin{center}
\fbox{\includegraphics[height=3cm]{validity_criterion}}
\caption{Validity criteria as an argument premise.}
\label{validity_criterion}
\end{center}
\end{figure}

To easily manage validity criteria, the relationships between validity criteria are also maintained in the argument graph using arguments such as "Published in journal with SJR score $>$ 0.67" $\rightarrow$ "Published in Annals of Mathematics" (validation criterion should be provided for such arguments too). By maintaining such relationships, Sophize allows its users to conveniently choose multiple sources they consider to be reliable with a single manually chosen criterion.

@@

A proposition is true if it is the conclusion of a valid argument and the premises of the argument are also true. Of course, this means that we need to start with some propositions that are assumed to be true without further justification. Traditionally, these are mathematical axioms and are considered to be self-evidently true. In mathematical inquiry, theorems are proven/disproven in one or more theories, each with its own set of axioms. Each theory may have its theoretical advantages and practical uses and there is no universal basis to choose one theory over another when trying to assert a proposition's truth/falsity.

Sophize allows any set of propositions to be considered true without evidence. Such a set of propositions is called a \textbf{belief set}. Propositions representing argument validity criteria are also part of belief sets. All propositions that are considered to be true in a belief set without proof are called its `unsupported propositions'. In addition, a belief set can also contain a number of `proof-generating machines' (e.g. for managing axiom-schemas) that will be discussed later. For ease of use, belief sets can also contain other belief sets.

The truth-value of a proposition is only defined within a belief set - there is no notion of absolute or universal truth on the Sophize platform. Sophize maintains proof graphs for all its belief sets.

@@

\subsection{Potential Advantages}

Logical organization of mathematics can lead to many advantages. We list some interesting ones below.

This knowledge organization scheme facilitates the re-use of propositions, and arguments across different theories. For example, to organize knowledge of different constant-curve geometries, we would create 4 belief sets - one each for absolute, Euclidean, hyperbolic and elliptic geometries. The latter three belief sets would comprise the absolute geometry belief set and the appropriate version of the parallel postulate needed for them. In such a scheme all the arguments (proofs) added to the argument graph for absolute geometry would automatically be shared by the other three geometries. The other three geometries need to only add the arguments that utilize their version of the parallel postulate.

Another use case would be to understand the consequences of important conjectures like the Reimann hypothesis by creating two belief sets - one where the conjecture is true and one where it is false. The proof graphs would systematically store such conjectural knowledge which often gets lost especially when results from multiple conjectures are involved. Complexity theory is one such field where this need is quite apparent. On a more practical note, several information security systems build on the conjectural developed with the assumption $P \neq NP$.

Proof graphs are currently organized as trees as seen in Fig \ref{proof_tree}. Users can keep exploring the proof till they reach a point where the propositions are already familiar to them. A future application of proof graphs would be to generate narrative proofs based on tailored to the reader's current understanding. This could be achieved in a classroom setting where the current level of mathematical understanding of students is known or assumed. It may also be possible to infer this from the user's activity on the platform.

\subsection{Limitations}
The scheme described in this section has a semantic notion of negation of a proposition to track contradictions. Thus it is incapable of organizing knowledge that depends on more complex calculi such as many-valued logics.

\section{Proof-Generating Machines (PGM)}
\label{sec:pgm}
A lot of mathematical knowledge is computed. Computation machines (e.g. calculators, algebras) can be seen as performing the following tasks
\begin{itemize}
\item Parsing the input into a language that can represent math formulae. 
\item Finding and outputting an equivalent form of the formula.
\end{itemize}
Thus, their focus is not on finding the proof but only finding the equivalent (simplified) representation of the input formula (Although many computer algebras do show the steps of the computation.). With Sophize, the focus instead is to generate proofs of the input propositions. A \textbf{proof-generating machine} (PGM) is a computer program that generates a partial proof graph against or in favour of an argument. PGMs can also have the capability to perform tasks 1 and 2, mentioned above.

@@

The Sophize platform itself merely forwards the request to the associated PGM and performs some checks on the output returned. Any http server that can process a proof request can be easily plugged in to Sophize platform using its HTTP address. One server can implement multiple PGMs. An overview of PGMs generating Metamath proofs is here: \url{https://youtu.be/hJtEIo3ioLM}

\subsection{`Active' PGMs}
As with other parts of Sophize, the validity of the output of a PGM is not taken for granted. Each PGM must specify all the propositions (premises) that it will use in the root nodes of the proof graph it returns. In addition, a PGM may also delegate parts of proofs to another PGM. These PGMs must also be specified. Similar to argument validation criterion, each PGM should also have premise indicating PGM's correctness criteria (e.g. tested by X, verified by some committee, another program etc.)

A PGM is considered `active' in a belief set iff:
\begin{enumerate}
\item all of its premises are true in the belief set.
\item all the PGMs it delegates-to are active in the belief set.
\end{enumerate}

The output of a PGM is not used in belief sets in which it is not active.

@@

\item Truth value, i.e., whether the PGM considers the input proposition (provided or computed) true or false. PGM can also indicate that it doesn't understand the input or can't prove/disprove the proposition.

@@

\begin{enumerate}
\item Arguments whose the premises are limited to the premises of the PGM.
\item Special `Arguments' that indicate that the generation of the proof of this argument's conclusion must be delegated to another PGM. To paginate large proofs, a PGM may also delegate to itself after generating some (say 100) arguments.
\end{enumerate}
\subsection{Materializing results}

@@

In the author's view, it is crucial to maintain the proofs graphs for building trustable knowledge. However, the presentation of proofs as graphs to users is perhaps not very convenient. Traditional text-based narration remains indispensable when presenting knowledge to the reader.

@@

With Sophize Markdown, we can easily add \LaTeX\space by enclosing it between two \$ signs (or \$\$ for display math). A detailed specification of Sophize Markdown is available \cite{EasyChair6207} and many of its features are demonstrated at \url{https://youtu.be/5UYOpQwcjCk}. We highlight relevant features here:

\subsection{Embedding Semantic Data}

With Sophize Markdown we can easily link resources such as terms, propositions, and arguments. Links can be added using the pound sign (\#) and the resource's URI. There are multiple resource link options that can be used for changing the behaviour and text of the resource link. The following example utilizes some of the options.

\subsubsection*{Sample Markdown Code}

@@

A conic section is a curve obtained as the intersection of the 
#(wiki/T_conical_surface, 'surface') of a #wiki/T_cone with a 
#wiki/T_plane. There are three types of conic sections.

@@

\paragraph{The above code is rendered as:}
\begin{mdframed}
\section*{Conic Section}
A conic section is a curve obtained as the intersection of the surface of a \dashuline{cone} with a \dashuline{plane}. There are three types of conic sections.

@@

In the above example, the `EXPAND' option expanded the definition of the term, and statement of the proposition in place. Thus we were able to easily re-use concepts and theorems already extracted from other sources (say, Wikipedia and the Oxford dictionary in this case). Clicking on the underlined concepts, pops-up a modal dialog with the definition of the term clicked.

@@

When a belief set is chosen, Sophize markdown adds an icon next to each proposition link. This icon indicates whether the proposition has been proven or dis-proven (or both) in the currently selected belief set using an appropriate symbol such as a check-mark or a cross. Clicking on the icon brings up the proof graph as shown in Fig.~\ref{proof_tree}. If the users switches from one belief set to another, the icon is updated based on the truth values in the new belief set but other parts of the article are unchanged. This functionality is demonstrated in the demo in subsection \ref{sec:sophize_proofs}. Similarly, an icon is added next to a proof generating machine which indicates whether the PGM is active or not.

\subsection{Formal Language Support}
In the case of formal languages, when propositions can be fully parsed, Sophize can automatically attach links to appropriate resources. This provides a convenient interface, where the user types in the native language, and the final output automatically allows users to explore all concepts that make up the input statement in depth. Currently, this is demonstrated in the use of Sophize Markdown with the Metamath language.

\section{Building a Modern Mathematics Library}
\label{sec:lib}
This sections discusses a some practical issues need to be overcome to build a modern scalable library with reliable information.

\subsection{Semantic Data Extraction}
Utilizing the full range of features offered by the Sophize platform requires the availability of semantic information. However, extracting semantic information from unstructured or semi-structured literature is a difficult problem. This problem can be seen as composed of several sub-problems:
\begin{itemize}
\item Finding definitions and propositions in prose.
\item Associating the appropriate entity with a term used in prose for quick lookup.
\item Identifying arguments along with premises and conclusion.
\item \label{prob:dup} Identifying duplicate definitions and propositions used in different sources.
\end{itemize}

Solving these problems would require manual effort (e.g. crowd-sourcing) combined with automated extraction tools. Recent progress in machine learning and natural language processing techniques \cite{ginev2019scientific} does make this problem seem tractable.

@@

Sophize divides its data into groups called datasets for copyright and data access management. For example, the 'wiki' dataset that extracts knowledge from Wikipedia mathematics glossaries releases its data under `CC-BY-SA' license, whereas the `metamath' dataset is public (`CC0'). Our signed-in users are free to create their own datasets and we allow them to choose who can read, write or comment in these datasets. Another interesting access control is the restriction of use of resources (especially argument-validity criteria) from an otherwise open-to-read dataset.

Information can be added to Sophize platform using intuitive web forms. However, it is infeasible to extract information from large repositories in this manner. Sophize publishes its data schema (explained in Section \ref{sec:core}) and helper libraries in multiple languages. The extracted information is can thus be stored as text (JSON) files in a repository. Sophize provides dataset owners the ability to sync knowledge such repositories to the central Sophize database.
